Make any stupid plot you want

### 2020-12-22 draft/sections/math.tex
sorting out notation + fleshing out

\begin{equation}
    A: O(E) \rightarrow O(H)
    \label{eq:artist}
\end{equation}

\begin{itemize}
\item $A$ is the function that converts an instance of data $\Gamma(E)$ to an instance of a visual representation $\Gamma(H)$ 
\item $E$ is a locally trivial fiber bundle over $K$ representing data space.
\item $K$ is a triangulizable space encoding the connectivity of the observations in the data. 
\item $H$ is a fiber bundle over $S$ representing visual space
\item $S$ is a simplacial complex of triangles encoding the connectivity of the visualization of the data in $E$
\item $\tau: K\rightarrow E$ is the data being visualized
\item $\rho: S \rightarrow H$ is the render map
\end{itemize}

@@

\begin{equation}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[r, "\pi"] & B
    \end{tikzcd}
\end{equation}

where there is a bijection from $F$ to every fiber $F_b$ over point $b \in B$ in $E$ and the function $\pi: E \rightarrow B$ is the map into the $B$ quotient space of $E$.

\subsubsection{Base space}
\begin{figure}[ht]
    \includegraphics[width=0.4\linewidth]{figures/sections/math/k_qspace.png}
    \label{fig:kquote}
\end{figure}

@@

\begin{equation}
    \begin{tikzcd}
        \pi^{-1}(U) \arrow[r, "\varphi"] \arrow[d, "\pi"'] & U \times F \arrow[ld, "\mathrm{proj}_U"] \\
        U                                                  &                                         
    \end{tikzcd}
    \label{eq:local_trivial}
\end{equation}
such that $\varphi: \pi^{-1}(U) \rightarrow U \times F$ is a homeomorphism where $\pi$ and $\mathrm{proj}_U$ both map to $U$ and the fiber over $k$ $F_b = \pi^{-1}({b \in B}) $ is homomorphic to the fiber $F$.

The section $f$ is the mapping from base space to total space $f: B\rightarrow E$ 
\begin{equation}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[d, "\pi"']           \\
        & B \arrow[u, "f"', bend right]
    \end{tikzcd}
\end{equation}

such that it is the right inverse of $\pi$
\begin{equation}
    \pi(f(b)) = b \text{ for all } b \in B 
\end{equation}

In a locally trivial fiber bundle, $E = B \times F$ \cite{rowlandFiberBundle,FiberBundle2020}:
\begin{equation}
f(b) = (b, g(b))
\end{equation}

@@

\subsubsection{Sheaf and Stalk}
As described in equation~\ref{eq:local_trivial}, there is a local space $U \subset B$ around every $b$. The inclusion map $\iota: U \rightarrow B$ can be pulled back such that $\iota^{*}E$ is the space of $E$ restricted over $U$. 
\begin{equation}
    \begin{tikzcd}
        \iota^*E \arrow[d, "\pi"']           & E \arrow[d, "\pi"'] \arrow[l, "\iota^*"']         \\
        U \arrow[u, "\iota^*f"', bend right] & B \arrow[u, "f"', bend right] \arrow[l, "\iota"']
        \end{tikzcd}
\end{equation}

The localized section of fibers $\iota^*f: U \rightarrow \iota^*E$ is the sheaf $O(E)$ with germ of $\xi^*f$. The neighborhood of points the sheaf lies over is the stalk $\mathscr{F}_b$ \cite{StalkSheaf2019,spanier1989algebraic}

\begin{equation}
    \iota^{-1}\mathscr{F}(\{b\}) = \varinjlim_{b\subseteq U}\mathscr{F}(U) =  \varinjlim_{b \in U} = \mathscr{F}_b 
\end{equation}

@@

\subsection{Data Model}

We use a fiber bundle model to represent the data, as proposed by Butler.  
\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}.
\begin{figure}[ht]
    \includegraphics[width=.2\linewidth]{figures/sections/math/fiberbundle.png}
    \caption{write up some words here}
    \label{fig:fiberbundle}
\end{figure}

@@

\subsubsection{Base Space}

\begin{figure}[ht]
    \includegraphics[width=0.2\linewidth]{figures/sections/math/temp_1k.png}
    %% add box around neighboring P and Map
    \label{fig:k_data}
\end{figure}

@@

\subsubsection{Fiber Space $F$}
\label{sec:data_fiber}
We use Spivak's formalization of data base schemas as the basis of our fiber space $F$ \cite{spivakSIMPLICIALDATABASES}. He defines the type specification 
\begin{equation}
\pi: U \rightarrow DT
\end{equation}

@@

\begin{figure}[ht]
    \includegraphics[width=0.2\linewidth]{figures/sections/math/temp_2f.png}
    \label{fig:}
\end{figure}
\begin{figure}[ht]
    \includegraphics[width=0.2\linewidth]{figures/sections/math/temp_3f.png}
\end{figure}

@@

Where the section $\rho: S \rightarrow H$
\begin{equation}
    \begin{tikzcd}
        D \arrow[r, hook] & H \arrow[d, "\pi"']              \\
                          & S \arrow[u, "\rho"', bend right]
        \end{tikzcd}
\end{equation}

 is mapping from a region $s$ on a mathematical encoding of the image to a region $xy$ on the screen that the renderer then maps to pixel space.

 \subsubsection{Section $\rho$}
 For a physical screen display, the predisplay space is a trivial fiber bundle $H=\mathbb{R}^{7}\times S$ such that $\rho$ is
\begin{equation}
    \rho(s)  = \{x, y, z, r, g, b, a\}
    \label{eq:rho}
\end{equation}

To draw an image, a region, $H$ is inverse mapped into a region $s \in S$ where
\begin{equation}
s = \rho^{-1}_{\tiny{XY}}(xy)
\end{equation}
such that the rest of the fields in $\mathbb{R}^{7}$ are then integrated over $s$ to yield the remaining fields:
%%change the integral to a double integral
\begin{align}
    r &= \iint\limits_s \rho_{\tiny{R}}(s)ds^{2}\\
    g &= \iint\limits_s \rho_{\tiny{G}}(s)ds^{2}\\
    b &= \iint\limits_s \rho_{\tiny{B}}(s)ds^{2}
\end{align}

@@

The artist is a mapping from the sheaf $O(E)$ representing the data to a pre-render space $O(H)$. 
\begin{equation}
    A: O(E) \rightarrow O(H)
\end{equation}
The artist is composed of two stages, data $E$ to visual mappings $V$ and $V$ mappings composited to a visual in prerender space $H$. The visual fiber bundle $V$ shares basespace $K$ with $E$:

@@

with fiber $P$ that encodes visual paramters such as position or color. As in section~\ref{sec:data_fiber}, the fields in $P$ are strongly typed. Inside the artist is also the equivalance class map $\xi: S \rightarrow K$ that goes from a region $s$ to its associated point $k$
\begin{equation}
    \begin{tikzcd}
        E \arrow[d, "\pi"'] & H \arrow[d, "\pi"'] \\
        K                   & S \arrow[l, "\xi"']
        \end{tikzcd}
\end{equation}
such that when $\xi(s) = k$, $\xi^*\tau(s) = \tau(k)$.

\subsubsection{Visual Mapping}

The function $\nu: E \rightarrow V$ maps from data space to visual space  \cite{carpendaleVisualRepresentationSemiology,bertinIIPropertiesGraphic2011,munznerWhatDataAbstraction2014} space
\begin{equation}
    \begin{tikzcd}
        E \arrow[d, "\pi"'] \arrow[r, "\nu"] & V \arrow[ld, "\pi"'] \\
        K                                    &                     
        \end{tikzcd}
\end{equation}

such that the visual section $\mu: K \rightarrow V$  is equivalent to the visual transformation of the data $\mu(k) = \nu\circ\tau(k)$
\begin{equation}
    \begin{tikzcd}
        E \arrow[r, "\nu"]                      & V \\
        K \arrow[ru, "\mu"'] \arrow[u, "\tau"'] &  
        \end{tikzcd}
\end{equation}

The constraint on $\nu: \tau\rightarrow \mu$ is that $\tau$ and $\mu$ must be equivariant.

%%set up proofs here

\subsubsection{Visual Composition}
The visual fiber bundle $V$ gets pulled back over $S$ via $\xi$ such that 
\begin{equation}
    \begin{tikzcd}
        \xi^*V \arrow[r, "Q"] \arrow[rd, "\xi^*\pi"'] & H \arrow[d, "\pi"] \\
                                                      & S                 
    \end{tikzcd}
\end{equation}
the function $Q:\xi^*V\rightarrow H$ composites points in $V$ into $\mathbb{R^7}$ tuples. The section $\mu$ is pulled over $s$

@@

such that the composition $Q$ of $\mu$ is equivalent to the render $\rho(s) = Q\circ\mu(s)$.

>>>>

\begin{equation}
    \label{eq:artist}
    A: \mathcal{O}(E) \rightarrow \mathcal{O}(H)
\end{equation}

\begin{itemize}
\item $A$ is the function that converts an instance of data $\Gamma(E)$ to an instance of a visual representation $\Gamma(H)$ 
\item $E$ is a locally trivial fiber bundle over $K$ representing data space.
\item $K$ is a triangulizable space encoding the connectivity of the points in $E$
\item $H$ is a fiber bundle over $S$ representing visual space
\item $S$ is a simplacial complex encoding the visualization
\end{itemize}

@@

\begin{equation}
    \label{eq:fiber_bundle}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[r, "\pi"] & B
    \end{tikzcd}
\end{equation}

where there is a bijection from $F$ to every fiber $F_b$ over point $b \in B$ in $E$ and the function $\pi: E \rightarrow B$ is the map into the $B$ quotient space of $E$. By defintion of a fiber bundle, $\pi$ is always a mapping from total space to base space, independent of the points $p \in E$, and therefore we call this mapping $\pi$ for all the fiber bundles in the model.

\subsubsection{Base space}
\begin{figure}[ht]
    \label{fig:base_space}
    \includegraphics[width=0.4\linewidth]{figures/sections/math/k_qspace.png}
\end{figure}

@@

\begin{equation}
    \label{eq:fiber_local_trivial}
    \begin{tikzcd}
        \pi^{-1}(U) \arrow[r, "\varphi"] \arrow[d, "\pi"'] & U \times F \arrow[ld, "\mathrm{proj}_U"] \\
        U                                                  &                                         
    \end{tikzcd}
\end{equation}
such that $\varphi: \pi^{-1}(U) \rightarrow U \times F$ is a homeomorphism where $\pi$ and $\mathrm{proj}_U$ both map to $U$ and the fiber over $k$ $F_b = \pi^{-1}({b \in B}) $ is homomorphic to the fiber $F$.

The section $f$ is the mapping from base space to total space $f: B\rightarrow E$ 
\begin{equation}
    \label{eq:section}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[d, "\pi"']           \\
        & B \arrow[u, "f"', bend right]
    \end{tikzcd}
\end{equation}

such that it is the right inverse of $\pi$
\begin{equation}
    \label{eq:section_domain}
    \pi(f(b)) = b \text{ for all } b \in B 
\end{equation}

In a locally trivial fiber bundle, $E = B \times F$ \cite{rowlandFiberBundle,FiberBundle2020}:
\begin{equation}
    \label{eq:section_return}
f(b) = (b, g(b))
\end{equation}

@@

\subsubsection{Sheaf and Stalk}
As described in equation~\ref{eq:local_trivial}, there is a local space $U \subset B$ around every $b$. The inclusion map $\iota: U \rightarrow B$ can be pulled back such that $\iota^{*}E$ is the space of $E$ restricted over $U$. 
\begin{equation}
    \label{eq:sheaf}
    \begin{tikzcd}
        \iota^*E \arrow[d, "\pi"']           & E \arrow[d, "\pi"'] \arrow[l, "\iota^*"']         \\
        U \arrow[u, "\iota^*f"', bend right] & B \arrow[u, "f"', bend right] \arrow[l, "\iota"']
        \end{tikzcd}
\end{equation}

The localized section of fibers $\iota^*f: U \rightarrow \iota^*E$ is the sheaf $\mathcal{O}(E)$ with germ of $\xi^*f$. The neighborhood of points the sheaf lies over is the stalk $\mathcal{F}_b$ \cite{StalkSheaf2019,spanier1989algebraic}

\begin{equation}
    \label{eq:stalk}
    \iota^{-1}\mathcal{F}(\{b\}) = \varinjlim_{b\subseteq U}\mathcal{F}(U) =  \varinjlim_{b \in U} = \mathcal{F}_b 
\end{equation}

@@

\subsection{Data Model}
As proposed by Butler \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, we model data as a fiber bundle:
\begin{equation}
    \label{eq:data_fiber_bundle}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[r, "\pi"] & B
    \end{tikzcd}
\end{equation}

with a total space $E$, fiber space $F$, and base space $K$ which can be thought of as a set of keys $k$. The section $\tau:K \rightarrow E$ is an instance of the data that lies in $E$ and is discussed in section~\ref{sec:data_fiber}.

\subsubsection{Example}

\begin{figure}[ht]
    \label{fig:data_fiber_bundle}
    \includegraphics[width=.2\linewidth]{figures/sections/math/fiberbundle.png}
    \caption{write up some words here}
\end{figure}

@@

\subsubsection{Base Space}
The base space $K$ is a representation of the connectivity of the data, specifically whether the points in $E$ are discrete or sampled from a continous space. The same dataset can be expressed with different $K$.

\subsubsection{Example}
\begin{figure}[ht]
    \includegraphics[width=0.2\linewidth]{figures/sections/math/temp_1k.png}
    %% add box around neighboring P and Map
    \label{fig:k_data}
\end{figure}

@@

\subsubsection{Fiber and Sections}
\label{sec:data_fiber}
We use Spivak's formalization of data base schemas as the basis of our fiber space $F$ \cite{spivakSIMPLICIALDATABASES}. He defines the type specification 
\begin{equation}
\pi: U \rightarrow DT
\end{equation}

@@

\subsubsection{Example}
\begin{figure}[ht]
    \includegraphics[width=0.2\linewidth]{figures/sections/math/temp_2f.png}
    \label{fig:}
\end{figure}
\begin{figure}[ht]
    \includegraphics[width=0.2\linewidth]{figures/sections/math/temp_3f.png}
\end{figure}

@@

\subsubsection{Section $\rho$}

A section $\rho: S \rightarrow H$
\begin{equation}
    \begin{tikzcd}
        D \arrow[r, hook] & H \arrow[d, "\pi"']              \\
                          & S \arrow[u, "\rho"', bend right]
        \end{tikzcd}
\end{equation}

 is a mapping from a region $s$ on a mathematical encoding of the image to a region $xy$ on the screen that the renderer then maps to visual space as defined in D. For a physical screen display, the predisplay space is a trivial fiber bundle $H=\mathbb{R}^{7}\times S$ such that $\rho$ is
\begin{equation}
    \rho(s)  = \{x, y, z, r, g, b, a\}
    \label{eq:rho}
\end{equation}

To draw an image, a region, $H$ is inverse mapped into a region $s \in S$ where
\begin{equation}
s = \rho^{-1}_{\tiny{XY}}(xy)
\end{equation}
such that the rest of the fields in $\mathbb{R}^{7}$ are then integrated over $s$ to yield the remaining fields:
\begin{align}
    r &= \iint\limits_s \rho_{\tiny{R}}(s)ds^{2}\\
    g &= \iint\limits_s \rho_{\tiny{G}}(s)ds^{2}\\
    b &= \iint\limits_s \rho_{\tiny{B}}(s)ds^{2}
\end{align}

@@

In this section we will define the artist as a mapping from the sheaf $O(E)$ representing the data to a pre-render space $O(H)$. 
\begin{equation}
    A: \mathcal{O}(E) \rightarrow \mathcal{O}(H)
\end{equation}
The artist decomposes to data $E$ to visual mappings $V$ and $V$ mappings composited to a visual in prerender space $H$. In order to define $A$, we first define the visual fiber bundle $V$ that shares basespace $K$ with $E$:

@@

with fiber $P$ that encodes visual parameters such as position or color. The section $\mu: V \rightarrow K$

\begin{equation}
    \begin{tikzcd}
        P \arrow[r, hook] & V \arrow[d, "\pi"']              \\
                          & K \arrow[u, "\mu"', bend right]
        \end{tikzcd}
\end{equation}

returns a record in $P$. The fiber $P$ is the space of visual parameters %% defined like P, dependent on Q, P is targeting our library

\subsubsection{Data Screen Mapping}
The data bases pace $K$ can be considered a subspace of the screen base space $S$ such that $\xi: S \rightarrow K$ is a deformation retraction
\begin{equation}
    \begin{tikzcd}
        E \arrow[d, "\pi"'] & H \arrow[d, "\pi"'] \\
        K                   & S \arrow[l, "\xi"']
        \end{tikzcd}
\end{equation}

that goes from a region $s \in S_{k}$ to its associated point $k$, such that when $\xi(s) = k$, $\xi^*\tau(s) = \tau(k)$.

\subsubsection{Visual Channels}

The function $\nu: E \rightarrow V$ is a mapping from the data bundle to the visual bundle 
\begin{equation}
    \begin{tikzcd}
        E \arrow[d, "\pi"'] \arrow[r, "\nu"] & V \arrow[ld, "\pi"'] \\
        K                                    &                     
        \end{tikzcd}
\end{equation}

such that a visual section $\mu: K \rightarrow V$  is equivalent to the visual transformation of the data $\mu(k) = \nu\circ\tau(k)$

%%throw out this diagram
%%something about pushforward ...
%% \mu is pushfowwards of \tau = \mu = \nu(lowercase * ,\tau), \nu: *\tau \mapsto \mu
%% \nu acts from subspace E to subspace V
The constraint on $\nu: \tau\rightarrow \mu$

%%set up proofs here, are x&y transfereble here? (x,y coupled together, Es bundled by what nu is promosing, maps over to V in indepdent ways 
%E ^ V decomposed in one to one correspondances, \nu_0 is map from E_0 to V_0...\nu_n from E_n, V_n...., \nu = {\nu_0....\nu_n}

\paragraph{Allowable Maps}
For example, stevens: preserve structure ....
Some examples of allowable maps are described in Bertin/Munzner:
nus are constrained to preserve..., expect functorial nus - transform on input side to transform on output side (E1-E1, V1-V1, nu) <- condition
that's how we're constraining nu. 
functorial in $\nu$ component

\subsubsection{Constructing Marks}
The visual fiber bundle $V$ gets pulled back over $S$ via $\xi$ such that 
\begin{equation}
    \begin{tikzcd}
        \xi^*V \arrow[r, "Q"] \arrow[rd, "\xi^*\pi"'] & H \arrow[d, "\pi"] \\
                                                      & S                 
    \end{tikzcd}
\end{equation}
the function $Q:\xi^*V\rightarrow H$ composites points in $V$ into $\mathbb{R^7}$ tuples. The section $\mu$ is pulled over $s$

@@

such that the composition $Q$ of $\mu$ is equivalent to the render $\rho(s) = Q\circ\mu(s)$. 
(In practice $\mu \approxeq \xi^*\mu$)
$Q$ is the constructor for the mark, 
one $Q$ per artist, A = sum of artists
$Q$ as there's default line, point, etc that's stretched and tweaked and whatever else by the $\mu$
visual variable + structure preserving, Q must have map that commutes such that a change in $\mu$ carries over somehow to $\rho$, functorial in its arguments - commutative diagram - into rho
